---
title: "Visualizing Large Distributions"
format:
  html:
    keep-md: true
editor: visual
---

# Task Description

In this task, we will explore and analyze the `nycflights13::flights` dataset using R. To begin, make sure you have the `tidyverse` and `nycflights13` packages installed and loaded.

```{r}
# Load the tidyverse package
library(tidyverse)

```

```{r}
# Load the nycflights13 package (if not already installed)
# install.packages("nycflights13")

# Load the flights dataset
library(nycflights13)
data(flights)

```

```{r}

?nycflights13::flights

```

# Summary Plot

I arrived at choosing the 'dep_delay' and 'arr_delay' columns then visualized the distribution for Departure Delay and Arrival Delay. In the insights of the Distribution Delay shown the data points closer to 0-1 min and same for Departure Arrival close to 0-1 min. For the 'Bivariate Summary of Departure Delay vs. Arrival Delay' see high cluster of smaller delay to smaller arrival times in minutes.

```{r}

variable1 <- "dep_delay"
variable2 <- "arr_delay"


dep_delay_plot <- ggplot(flights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Departure Delay",
       x = "Departure Delay (minutes)",
       y = "Frequency")

arr_delay_plot <- ggplot(flights, aes(x = arr_delay)) +
  geom_histogram(binwidth = 10, fill = "salmon", color = "black") +
  labs(title = "Distribution of Arrival Delay",
       x = "Arrival Delay (minutes)",
       y = "Frequency")

bivariate_summary_plot <- ggplot(flights, aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Bivariate Summary of Departure Delay vs. Arrival Delay",
       x = "Departure Delay (minutes)",
       y = "Arrival Delay (minutes)")


dep_delay_plot
arr_delay_plot
bivariate_summary_plot


```
