---
title: "Mean Heightst of Countries Over Time"
author: "David Pineda"
format:
  html:
    keep-md: true
editor: visual
---

# Import Packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(forcats)
library(pander)

# Install and load the readxl package if you haven't already
# install.packages("readxl")
library(readxl)
```

```{r}

excel_file <- "/Users/daviedev/DS350_FA23_Pineda_Davi/week_04/Height.xlsx"

data <- read_excel(excel_file)

```

```{r}
colnames(data)[colnames(data) == 'Continent, Region, Country'] <- 'country'

```

```{r}
pivot_table <- data %>%
  pivot_longer(
    cols = -c(Code, country),
    values_to = "height.cm",
    names_to = "year_decade"
  )

```

```{r}
pivot_table <- pivot_table %>% 
  mutate(year_decade = sub("\\.0$", "", year_decade)) %>% 
  mutate(height.cm = round(height.cm, 1)) %>% 
  select(Code, country, height.cm, year_decade)
  
```

```{r}
sorted_pivot_table <- pivot_table %>%
  arrange(year_decade) %>% 
  drop_na()

sorted_pivot_table 
```

```{r}

df <- sorted_pivot_table  %>%
  mutate(height.inches = height.cm / 2.54)
```

```{r}

plot <- ggplot(df, aes(x = year_decade, y = height.inches),color= country) +
  geom_point(color="black", size=4) +
  geom_point(data=subset(df, country=="Germany"),color="red",size = 4) +
  labs(x = "Year Decade", y = "Height (inches)", title = "Height Trends by Country") +
  theme_minimal() 
  #scale_color_manual(values = c("other" = "black","Germany" = "red")) 
  #guides(color = guide_legend(order = 1))

plot
```

# Summary

Based on the available data, it does appear that humans have been getting taller over the years. The data, particularly the trend observed in Germany's growth from 1810 to 1980, suggests a consistent increase in height. It's important to note that this conclusion is based on the specific data set and time frame analyzed, and a comprehensive study would require a more extensive examination of global trends in human height over a longer period.
